---
title: "Lottery Leaflet Maps"
author: "Shreya Vuttaluru"
date: "9/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This markdown file maps census variables across census tracts in various states. For now, we're looking at just total population in key states identified by research, but this can be expanded to any state in turn by four variables: from ACS, poverty/education/income; from 2020 data, race (mapped by non-white percentage, but details on hover for each demographic breakdown). By Shreya Vuttaluru and Rachel Logan

```{r libraries, results=FALSE}
library(tidyverse)
library(janitor)
library(censusxy)
library(tidycensus)
library(dplyr)
#install.packages("leaflet")
library(leaflet)
library(sf)

### CENSUS API KEY
census_api_key("549950d36c22ff16455fe196bbbd01d63cfbe6cf")
```

##Getting Decennial geometries
```{r results=FALSE}
state_list <- fips_codes %>%
            distinct(state) %>%
            as.list()

state_list <-state_list$state[1:51]

decennial_vars <- load_variables(2020, dataset="pl")

nation_geometries <- get_decennial(year=2020, geography="tract",state=state_list, variables = "P3_001N", sumfile = "pl", geometry = TRUE)

clean_dec_nation_geometries <- nation_geometries %>% 
  clean_names() %>% 
  separate(name, into = c("tract","county","state"), sep=", ") %>% 
  mutate(total_pop=value) %>% 
  select(-variable,-value)

write_rds(clean_dec_nation_geometries,"../../data/processed/analysis/make_leaflet_maps/dec_nation_geometries.rds")
```


## Loading retailer information 
``` {r results=FALSE}
all_retailers <- read_rds("../../data/processed/a_retailers_geospatial/03_join_retailers/all_geocoded_retailers.rds")

retailers_w_tracts <- all_retailers %>%
  filter(!is.na(decennial_2020_tract_geoid))

join_retailers_populations <- retailers_w_tracts  %>% 
  mutate(GEOID = decennial_2020_tract_geoid) %>%
  left_join(nation_geometries)

## this df below has the number of retailers in a given tract w/ geometries
retailers_in_tracts <- join_retailers_populations %>%
  clean_names() %>% 
  separate(name, into = c("tract","county","state"), sep=", ") %>% 
  mutate(total_pop=value) %>%
  group_by(geoid, tract, county, state, total_pop) %>%
  summarize (
    count = n()
    ) %>%
  rename(retailer_count = count) %>%
  ungroup() 

retailers_in_tracts <- clean_dec_nation_geometries %>%
  inner_join(retailers_in_tracts) %>%
  ## some tracts have no people, but a lot of retailers, and vice versa. filter helps. 
  filter(total_pop > 100) %>%
  mutate( 
    retailer_concentration = (retailer_count / total_pop) * 10000)  %>%
  ungroup()
  
write_rds(retailers_in_tracts,"../../data/processed/analysis/make_leaflet_maps/retailers_in_tracts.rds")
```

##Getting ACS geometries
```{r results=FALSE}
acs_nation_geometries <- get_acs(geography = "tract", state=state_list, variables = "B01001_001", geometry = TRUE)

clean_acs_nation_geometries <- acs_nation_geometries %>% 
  clean_names() %>% 
  separate(name, into = c("tract","county","state"), sep=", ")%>% 
  mutate(total_pop=estimate) %>% 
  select(-variable,-estimate,-moe)

write_rds(clean_acs_nation_geometries,"../../data/processed/analysis/make_leaflet_maps/acs_nation_geometries.rds")
```

## Preparing leaflet
```{r}
MD_dec_geometries <- clean_dec_nation_geometries %>% 
  filter(state== "Maryland")

pal <- colorNumeric(
  palette = "inferno",
  domain = MD_dec_geometries$total_pop
)

leaflet() %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(data = MD_dec_geometries,
              color = ~pal(total_pop),
              weight = 0.5,
              smoothFactor = 0.2,
              fillOpacity = 0.75,
              label = ~total_pop) %>%
  addLegend(
    position = "bottomright",
    pal = pal,
    values = MD_dec_geometries$total_pop,
    title = "total pop <br>per census tract"
  )
```

## Retailer Concentration Mapping: Maryland
```{r}
#clean_dec_nation_geometries <- readRDS("dec_nation_geometries.rds")

dec_geometries <- retailers_in_tracts %>% 
  filter(state== 'Maryland')

pal <- colorNumeric(
  palette = "plasma",
  domain = dec_geometries$retailer_concentration
)

leaflet() %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(data = dec_geometries,
              color = ~pal(retailer_concentration),
              weight = 0.5,
              smoothFactor = 0.2,
              fillOpacity = 0.75,
              label = ~retailer_concentration) %>%
  addLegend(
    position = "bottomright",
    pal = pal,
    values = dec_geometries$retailer_concentration,
    title = "retailer concentration<br>per 10,000 people"
  )
```


## Retailer Concentration Mapping: DC
```{r}
DC_dec_geometries <- retailers_in_tracts %>% 
  filter(state== "District of Columbia")

pal <- colorNumeric(
  palette = "plasma",
  domain = DC_dec_geometries$retailer_concentration
)

leaflet() %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(data = DC_dec_geometries,
              color = ~pal(retailer_concentration),
              weight = 0.5,
              smoothFactor = 0.2,
              fillOpacity = 0.75,
              label = ~retailer_concentration) %>%
  addLegend(
    position = "bottomright",
    pal = pal,
    values = DC_dec_geometries$retailer_concentration,
    title = "retailer concentration<br>per 10,000 people"
  )
```


## Retailer Concentration Mapping: Virginia
``` {r}
VA_dec_geometries <- retailers_in_tracts %>% 
  filter(state== "Virginia")

pal <- colorNumeric(
  palette = "plasma",
  domain = VA_dec_geometries$retailer_concentration
)

leaflet() %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(data = VA_dec_geometries,
              color = ~pal(retailer_concentration),
              weight = 0.5,
              smoothFactor = 0.2,
              fillOpacity = 0.75,
              label = ~retailer_concentration) %>%
  addLegend(
    position = "bottomright",
    pal = pal,
    values = VA_dec_geometries$retailer_concentration,
    title = "retailer concentration<br>per 10,000 people"
  )

```

## Retailer Concentration Mapping: Michigan
``` {r}
MI_dec_geometries <- retailers_in_tracts %>% 
  filter(state== "Michigan")

pal <- colorNumeric(
  palette = "plasma",
  domain = MI_dec_geometries$retailer_concentration
)

leaflet() %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(data = MI_dec_geometries,
              color = ~pal(retailer_concentration),
              weight = 0.5,
              smoothFactor = 0.2,
              fillOpacity = 0.75,
              label = ~retailer_concentration) %>%
  addLegend(
    position = "bottomright",
    pal = pal,
    values = MI_dec_geometries$retailer_concentration,
    title = "retailer concentration<br>per 10,000 people"
  )
```

## Retailer Concentration Mapping: Connecticut
``` {r}

CT_dec_geometries <- retailers_in_tracts %>% 
  filter(state== "Connecticut")

pal <- colorNumeric(
  palette = "plasma",
  domain = CT_dec_geometries$retailer_concentration
)

leaflet() %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(data = CT_dec_geometries,
              color = ~pal(retailer_concentration),
              weight = 0.5,
              smoothFactor = 0.2,
              fillOpacity = 0.75,
              label = ~retailer_concentration) %>%
  addLegend(
    position = "bottomright",
    pal = pal,
    values = CT_dec_geometries$retailer_concentration,
    title = "retailer concentration<br>per 10,000 people"
  )
```

## Retailer Concentration Mapping: South Carolina
``` {r}
SC_dec_geometries <- retailers_in_tracts %>% 
  filter(state== "South Carolina")

pal <- colorNumeric(
  palette = "plasma",
  domain = SC_dec_geometries$retailer_concentration
)

leaflet() %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(data = SC_dec_geometries,
              color = ~pal(retailer_concentration),
              weight = 0.5,
              smoothFactor = 0.2,
              fillOpacity = 0.75,
              label = ~retailer_concentration) %>%
  addLegend(
    position = "bottomright",
    pal = pal,
    values = SC_dec_geometries$retailer_concentration,
    title = "retailer concentration<br>per 10,000 people"
  )

```

## Retailer Concentration Mapping: New York
``` {r}

NY_dec_geometries <- retailers_in_tracts %>% 
  filter(state== "New York")

pal <- colorNumeric(
  palette = "plasma",
  domain = NY_dec_geometries$retailer_concentration
)

leaflet() %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(data = NY_dec_geometries,
              color = ~pal(retailer_concentration),
              weight = 0.5,
              smoothFactor = 0.2,
              fillOpacity = 0.75,
              label = ~retailer_concentration) %>%
  addLegend(
    position = "bottomright",
    pal = pal,
    values = NY_dec_geometries$retailer_concentration,
    title = "retailer concentration<br>per 10,000 people"
  )

```

